Data prep for Shiny app

### 2017-04-14 shiny_am_iucn/data_prep.Rmd
added coral maps functionality and lots of explanatory text

dir_git   <- '~/github/IUCN-AquaMaps'
dir_shiny <- file.path(dir_git, 'shiny_am_iucn')

@@

@@

Get coral species file and bring it over here; clean it up a bit for easier processing.

@@

write_csv(iucn_coral_fixed, file.path(dir_shiny, 'data', 'coral_spp_areas_app.csv'))

>>>>

dir_git   <- '~/github/IUCN-AquaMaps' %>% path.expand()
dir_shiny <- file.path(dir_git, 'shiny_am_iucn')

@@

## Get FAO regions

``` {r copy_fao_rgns}

# fao_files <- list.files(file.path(dir_git, 'clip_fao/fao_rgns'), pattern = 'fao_rgns')
# 
# file.copy(from = file.path(dir_git, 'clip_fao/fao_rgns', fao_files),
#           to   = file.path(dir_shiny, 'data', fao_files %>%
#                              str_replace('\\.', '_app.')))
# fao_rgn <- rgdal::readOGR(dsn = 'data', layer = "fao_rgns2_app") %>%
#   crop(extent(-180, 180, -90, 90))
# rgdal::writeOGR(fao_rgn, dsn = 'data', layer = "fao_rgns2_app", driver = 'ESRI Shapefile', overwrite.layer = TRUE)

@@

Get coral species alignment file and bring it over here; clean it up a bit for easier processing.

@@

write_csv(iucn_coral_fixed, file.path(dir_shiny, 'data', 'coral_align_app.csv'))

### Get coral cells and info for mapping

``` {r get_coral_info}

### load iucn spp_cells from main analysis; clip to just those in app
spp_list <- read_csv(file.path(dir_shiny, 'data', 'spp_list_app_stable.csv'))

coral_cells_app <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
  filter(iucn_sid %in% spp_list$iucn_sid)
# length(unique(coral_cells_app$iucn_sid))
write_csv(coral_cells_app, file.path(dir_shiny, 'data', 'coral_cells_app.csv'))

coral_spp_list  <- spp_list %>%
  select(iucn_sid, sciname) %>%
  filter(iucn_sid %in% coral_cells_app$iucn_sid) %>%
  distinct()
write_csv(spp_coral_maps, file.path(dir_shiny, 'data', 'coral_spp_list.csv'))

coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quad_sum.csv'))
write_csv(coral_quad_sum, file.path(dir_shiny, 'data', 'coral_quads_app.csv'))

``` {r simplify_bathymetry}
library(maptools)
library(tmap)

bathy_shp_orig <- rgdal::readOGR(dsn = file.path(dir_git, 'clip_depth/ne_10m_bathymetry_K_200'),
                            layer = 'ne_10m_bathymetry_K_200') %>%
  rgeos::gUnaryUnion()

keep_df <- data.frame(
  id   = 1:length(bathy_shp_orig@polygons[[1]]@Polygons),
  area = lapply(bathy_shp_orig@polygons, function(x) sapply(x@Polygons, function(y) y@area)) %>%
    unlist(),
  hole = lapply(bathy_shp_orig@polygons, function(x) sapply(x@Polygons, function(y) y@hole)) %>%
    unlist()
)
quantile(keep_df$area)
#           0%          25%          50%          75%         100% 
# 2.319330e-10 2.026033e-02 3.717621e-02 9.453268e-02 4.108813e+04

### ditch the small polys.  Tolerance? 6e-2 (.06) is .25° x .25°; can go bigger,
### perhaps up to a single .5° cell, or .25?
hole_thresh <- .0625
feat_thresh <- 2

keep_df1 <- keep_df %>%
  mutate(keep = TRUE,
         # keep = ifelse(area < hole_thresh, FALSE, keep))
         keep = ifelse(hole == TRUE  & area < hole_thresh, FALSE, keep),
         keep = ifelse(hole == FALSE & area < feat_thresh, FALSE, keep))

keep_ids <- keep_df1$id[keep_df1$keep]

bathy_shp <- bathy_shp_orig

bathy_shp@polygons[[1]]@Polygons <- bathy_shp@polygons[[1]]@Polygons[keep_ids]
bathy_shp@polygons[[1]]@plotOrder <- 1:length(bathy_shp@polygons[[1]]@Polygons)
bathy_shp@plotOrder <- 1:length(bathy_shp@polygons)

### need to repair comments... ugh
slot(bathy_shp, "polygons") <- lapply(slot(bathy_shp, "polygons"), checkPolygonsHoles)

# # will assign comments, seen by: 
# lapply(slot(bathy_shp, "polygons"), comment) 
# 
# # or remove them by: 
# slot(bathy_shp, "polygons") <- lapply(slot(bathy_shp, "polygons"), 
#    "comment<-", NULL)

### convert to spatial polygons dataframe
bathy_shp1 <- bathy_shp %>%
  SpatialPolygonsDataFrame(data = data.frame('depth' = 200))

### check new map against original map
tm_shape(bathy_shp_orig) + 
  tm_polygons(col = 'blue', lwd = 0) + 
  tm_shape(bathy_shp1) + 
  tm_polygons(col = 'red', 
              border.col = 'red3', 
              lwd = 0, alpha = .8)

writePolyShape(bathy_shp1,
               file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200/bathy_app_simple'))
file.copy(from = file.path(dir_git, 'clip_depth/ne_10m_bathymetry_K_200/ne_10m_bathymetry_K_200.prj'),
          to   = file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200/bathy_app_simple.prj'),
          overwrite = FALSE)
# rgdal::writeOGR(bathy_shp,
#                 dsn    = path.expand(file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200')),
#                 layer  = 'bathy_app_simple',
#                 driver = 'ESRI Shapefile',
#                 overwrite_layer = TRUE)
